simulation

### 2025-12-30 simul.Rmd
changed get_rho

## R Markdown

```{r}
# ==============================================================================
# EQUIVALENCE CHECK: exact_inprod() vs get_rho()
# ==============================================================================

library(fda)

# 1. DEFINE FUNCTIONS (To ensure environment is self-contained)
# ------------------------------------------------------------------------------

# A. Your Function (Logic: Matrix Multiplication)

# 2. PREPARE DATA
# ------------------------------------------------------------------------------
set.seed(123)
n_sample <- 50
n_basis  <- 10
basis <- create.bspline.basis(c(0,1), n_basis)

# Generate Random W (Data)
coefs_W <- matrix(rnorm(n_sample * n_basis), n_basis, n_sample)
Z_W <- matrix(rnorm(n_sample * 5), n_sample, 5)
fd_W <- fd(coefs_W, basis)
W_current <- list(
  Z = Z_W,
  functional_list = list(fd_W), # Single functional predictor for simplicity
  n_sample = n_sample,
  n_functional = 1,
  n_scalar = 5,
  n_basis_list = list(n_basis)
)

# CRITICAL STEP: Populate 'gram_list' for get_rho to work
# Your function assumes this exists. We must calculate it exactly as FDA does.
W_current$gram_list <- list(
  fda::inprod(basis, basis)
)

# Generate Random xi (Direction)
coefs_xi <- matrix(rnorm(n_basis), n_basis, 1) # 1 direction
Z_xi <- matrix(rnorm(5), 1, 5)
fd_xi <- fd(coefs_xi, basis)
xi_current <- list(
  Z = Z_xi,
  functional_list = list(fd_xi)
)

# 3. CONVERT INPUTS
# ------------------------------------------------------------------------------
# get_rho expects a flattened vector 'd_vec'.
# We must flatten xi_current: [functional_coefs | scalar_weights]
d_vec <- c(as.vector(coefs_xi), as.vector(Z_xi))

# 4. RUN COMPARISON
# ------------------------------------------------------------------------------

# Run My Function
val_exact <- exact_inprod(W_current, xi_current)

# Run Your Function (Note: d_vec is 1st arg, W is 2nd arg)
val_yours <- get_rho(d_vec, W_current)

# 5. VERDICT
# ------------------------------------------------------------------------------
diff <- max(abs(val_exact - val_yours))

cat("========================================\n")
cat("      EQUIVALENCE CHECK RESULT       \n")
cat("========================================\n")
cat(sprintf("Max Difference: %.10e\n", diff))

if(isTRUE(all.equal(val_exact, val_yours))) {
  cat("CONCLUSION: ✅ FUNCTIONS ARE IDENTICAL.\n")
  cat("Your implementation uses valid exact matrix algebra.\n")
} else {
  cat("CONCLUSION: ❌ FUNCTIONS DIFFER.\n")
}
```

```{r}
#' Exact Hybrid Inner Product (Matrix Based)
#' 
#' Calculates <W, xi> using Basis Inner Product Matrices (J).
#' Guarantees mathematical exactness for B-splines.
#' 
#' @param W Hybrid predictor (N samples)
#' @param xi Hybrid direction (1 sample/direction)
#' @return Numeric vector of length N (The scores)
exact_inprod <- function(W, xi) {

  # 1. Scalar Part (Dot Product)
  # (N x p) %*% (p x 1)
  scores <- as.matrix(W$Z) %*% t(as.matrix(xi$Z))

  # 2. Functional Part (Exact Integral via J matrix)
  for(k in seq_along(W$functional_list)) {
    fd_W  <- W$functional_list[[k]]
    fd_xi <- xi$functional_list[[k]]

    # Basis Inner Product Matrix: J[i,j] = Int(phi_i * phi_j)
    J <- fda::inprod(fd_W$basis, fd_W$basis)

    # Score = Coef_W^T * J * Coef_xi
    # (N x K) * (K x K) * (K x 1) -> (N x 1)

    # Efficient calculation:
    # We broadcast the (J %*% Coef_xi) vector across all N samples
    weighted_xi <- J %*% fd_xi$coefs

    # Add to scores (t(coefs) is N x K)
    scores <- scores + t(fd_W$coefs) %*% weighted_xi
  }

  return(as.vector(scores))
}
```

@@

  # 2. Get Scores (Rho) -- EXACT
  #rho_current <- exact_inprod(W_current, xi_current)
  rho_current <- get_rho(W_current, xi_current)

@@

ggplot(df_long, aes(x=factor(Iteration), y=value, fill=name)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(trans="log10") +
  labs(title="Annihilation Torture Test",
       subtitle="Even with high collinearity and penalties, leakage must be ~0",
       y="Magnitude (Log10)",
       x="Iteration") +
  theme_minimal() +
  scale_fill_manual(values=c("firebrick", "steelblue")) +
  geom_hline(yintercept=1e-10, linetype="dashed")
```

```{r}
# ==============================================================================
# STUDY: PLS Geometric Properties (Orthogonality & Relevance)
# ==============================================================================

@@

# ------------------------------------------------------------------------------
# 1. SETUP
# ------------------------------------------------------------------------------

lambda_scenarios <- list(
  "Scenario A: weak"    = c(0.1, 0.1),         # Pure NIPALS
  "Scenario B: mixed"  = c(0.1, 5),    # Mixed Smoothness
  "Scenario C: strong"   = c(5, 5)        # Strong Regularization
)

n_rep <- 100       # Repetitions
n_comp <- 6        # Components
results_ortho <- data.frame() # Store orthogonality results
results_corr  <- data.frame() # Store correlation results

# Data Generation Function
generate_robust_data <- function(n_sample = 200, n_scalar = 50) {
  t1 <- rnorm(n_sample, sd = 10)
  t2 <- rnorm(n_sample, sd = 5)
  t3 <- rnorm(n_sample, sd = 2)

  eval_points <- seq(0, 1, length.out = 100)
  basis <- create.bspline.basis(c(0, 1), 10)

  X1 <- outer(t1, sin(2*pi*eval_points)) + outer(t2, cos(2*pi*eval_points))
  fd1 <- Data2fd(eval_points, t(X1), basis)

  X2 <- outer(t1, sin(4*pi*eval_points)) + outer(t2, cos(4*pi*eval_points))
  fd2 <- Data2fd(eval_points, t(X2), basis)

  Z <- matrix(0, n_sample, n_scalar)
  for(j in 1:n_scalar) Z[,j] <- t1*runif(1) + t3*runif(1) + rnorm(n_sample)

  y <- t1 * 2 - t2 * 1.5 + rnorm(n_sample)
  W <- predictor_hybrid(Z, list(fd1, fd2), eval_points)

  return(list(W = W, y = y))
}

# Helper: Compute Inner Product of Two Hybrid Vectors
# (Required to check orthogonality of xi)
hybrid_inner_product <- function(h1, h2) {
  # Scalar dot product
  val <- sum(h1$Z * h2$Z)
  # Functional inner products
  for(k in seq_along(h1$functional_list)) {
    val <- val + inprod(h1$functional_list[[k]], h2$functional_list[[k]])
  }
  return(as.numeric(val))
}

# ------------------------------------------------------------------------------
# 2. RUN SIMULATION LOOP
# ------------------------------------------------------------------------------
message("Starting Extended Simulation...")
set.seed(999)

for (scenario_name in names(lambda_scenarios)) {
  lam_pair <- lambda_scenarios[[scenario_name]]

  for (i in 1:n_rep) {
    # Generate & Fit
    data <- generate_robust_data(n_sample = 200)
    processed <- split_and_normalize.all(data$W, data$y, train_ratio = 0.8)

    fit <- fit.hybridPLS(
      W = processed$predictor_train, 
      y = processed$response_train, 
      n_iter = n_comp, 
      lambda = lam_pair
    )

    # --- METRIC 1: SCORE (Rho) ORTHOGONALITY ---
    rho_mat <- do.call(cbind, fit$rho)
    cor_rho <- cor(rho_mat)
    diag(cor_rho) <- 0
    max_rho_overlap <- max(abs(cor_rho))

    # --- METRIC 2: DIRECTION (Xi) ORTHOGONALITY ---
    # We calculate the max pairwise inner product between different weight vectors
    # Note: xi are normalized, so self-inner-product is 1.
    xi_list <- fit$xi
    xi_overlaps <- numeric()

    for(j in 1:(n_comp-1)) {
      for(k in (j+1):n_comp) {
        # Dot product between direction j and direction k
        dp <- hybrid_inner_product(xi_list[[j]], xi_list[[k]])
        xi_overlaps <- c(xi_overlaps, abs(dp))
      }
    }
    max_xi_overlap <- max(xi_overlaps)

    # Record Orthogonality Stats
    results_ortho <- rbind(results_ortho, data.frame(
      Scenario = scenario_name,
      Repetition = i,
      Max_Score_Corr = max_rho_overlap,    # Should be ~0
      Max_Direction_Cos = max_xi_overlap   # Can be > 0
    ))

    # --- METRIC 3: SCORE vs RESPONSE CORRELATION ---
    # Does Comp 1 explain Y better than Comp 2?
    y_vec <- processed$response_train
    cor_y <- cor(rho_mat, y_vec)

    # Store trajectory for this repetition
    for(k in 1:n_comp) {
      results_corr <- rbind(results_corr, data.frame(
        Scenario = scenario_name,
        Repetition = i,
        Component = k,
        Correlation_with_Y = abs(cor_y[k, 1])
      ))
    }
  }
}

# ------------------------------------------------------------------------------
# 3. REPORT GENERATION
# ------------------------------------------------------------------------------

# --- TABLE 1: Orthogonality (Scores vs Directions) ---
summary_ortho <- results_ortho %>%
  group_by(Scenario) %>%
  summarise(
    Score_Orthogonality = mean(Max_Score_Corr),
    Direction_Overlap   = mean(Max_Direction_Cos)
  )

cat("\n=================================================================\n")
cat(" TABLE 1: Geometric Properties (Scores vs Directions)\n")
cat("=================================================================\n")
cat("NOTE: 'Score' should be 0 (Orthogonal). 'Direction' can be > 0.\n\n")
print(summary_ortho)

# --- VISUALIZATION: Relevance Decay ---
# We want to see if correlation drops as component index increases

summary_corr <- results_corr %>%
  group_by(Scenario, Component) %>%
  summarise(
    Mean_Corr = mean(Correlation_with_Y),
    SD_Corr = sd(Correlation_with_Y)
  )

p_decay <- ggplot(summary_corr, aes(x = Component, y = Mean_Corr, color = Scenario)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = Mean_Corr - SD_Corr, ymax = Mean_Corr + SD_Corr), width = 0.2) +
  labs(title = "Predictive Relevance Decay",
       subtitle = "Correlation between PLS Scores and Response Y",
       y = "|Correlation(Score, Y)|",
       x = "Component Index") +
  theme_bw() +
  scale_x_continuous(breaks = 1:n_comp)

print(p_decay)
```

```{r}
# ==============================================================================
# PLS VERIFICATION: Prediction & Beta Recovery
# ==============================================================================

library(fda)
library(ggplot2)
library(gridExtra)
library(FSHybridPLS) # Your package

# ------------------------------------------------------------------------------
# 1. GENERATE TRUTH (Linear Model)
# ------------------------------------------------------------------------------
generate_regression_truth <- function(n_sample = 200, n_eval = 100, noise_sd = 0.5) {

  eval_points <- seq(0, 1, length.out = n_eval)
  basis_obj <- create.bspline.basis(c(0, 1), 20)

  # --- A. True Coefficient Function (Beta) ---
  # A distinct shape: Sine wave with a trend
  true_beta_vec <- sin(2 * pi * eval_points) + 1.5 * eval_points
  # Center it (standard practice)
  true_beta_vec <- true_beta_vec - mean(true_beta_vec)

  # --- B. Functional Predictors (X) ---
  # Smooth random curves
  coefs <- matrix(rnorm(20 * n_sample), 20, n_sample)
  # Smooth the coefficients to make curves realistic
  smooth_mat <- chol(ar1_cov(20, 0.8))
  coefs <- t(smooth_mat) %*% coefs

  fd_obj <- fd(coefs, basis_obj)
  X_vals <- eval.fd(eval_points, fd_obj) # 100 x 200

  # --- C. Response (Y) ---
  # Y = Integral( X(t) * Beta(t) ) + Noise
  # Approx integral as dot product * dt
  dt <- 1 / n_eval

  # Calculate signal
  # (100 x 200) * (100 x 1) -> broadcasting requires care or loop
  # Signal = colSums( X * beta ) * dt
  signal <- colSums(X_vals * true_beta_vec) * dt

  # Add noise
  y <- signal + rnorm(n_sample, sd = noise_sd)

  # Scalar Z (Noise only, to prove PLS ignores it)
  Z <- matrix(rnorm(n_sample), n_sample, 1)

  W <- predictor_hybrid(Z, list(fd_obj), eval_points)

  return(list(
    W = W, y = y, 
    true_beta = true_beta_vec,
    eval_points = eval_points
  ))
}

# ------------------------------------------------------------------------------
# 2. RUN SIMULATION
# ------------------------------------------------------------------------------
set.seed(42)
# Generate
sim_data <- generate_regression_truth(n_sample = 200, noise_sd = 0.1)

# Split (Train/Test)
processed <- split_and_normalize.all(sim_data$W, sim_data$y, train_ratio = 0.7)

# Fit PLS
# Using 4 components (should be enough to capture the beta shape)
# Lambda = 0 ensures we test the core algorithm, not the penalty
fit <- fit.hybridPLS(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 4, 
  lambda = c(0),
  validation_data = list(W_test = processed$predictor_test, y_test = processed$response_test)
)

# ------------------------------------------------------------------------------
# 3. VERIFICATION PLOTS
# ------------------------------------------------------------------------------

# --- A. Prediction Check (Test Set) ---
# Reconstruct Y_pred for Test Set using the Final Beta (Comp 4)
beta_final <- fit$beta[[4]]
y_pred <- inprod.predictor_hybrid(processed$predictor_test, beta_final)
y_true <- processed$response_test

r2_test <- 1 - sum((y_true - y_pred)^2) / sum((y_true - mean(y_true))^2)

df_pred <- data.frame(Observed = y_true, Predicted = y_pred)

p1 <- ggplot(df_pred, aes(x = Observed, y = Predicted)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = paste0("Prediction Accuracy (R2 = ", round(r2_test, 3), ")"),
       subtitle = "Points should lie on the red diagonal") +
  theme_bw()

# --- B. Beta Recovery Check ---
# Extract estimated functional beta
beta_est_fd <- fit$beta[[4]]$functional_list[[1]]
beta_est_vec <- eval.fd(sim_data$eval_points, beta_est_fd)

# Normalize both to compare SHAPE (scale might differ due to normalization)
scale_vec <- function(x) as.vector((x - mean(x)) / sd(x))

df_beta <- data.frame(
  Time = sim_data$eval_points,
  True = scale_vec(sim_data$true_beta),
  Estimated = scale_vec(beta_est_vec)
)

beta_corr <- cor(df_beta$True, df_beta$Estimated)

p2 <- ggplot(df_beta, aes(x = Time)) +
  geom_line(aes(y = True, color = "True Beta"), linewidth = 1.2, linetype = "dashed") +
  geom_line(aes(y = Estimated, color = "PLS Estimate"), linewidth = 1) +
  scale_color_manual(values = c("True Beta" = "black", "PLS Estimate" = "red")) +
  labs(title = paste0("Beta Reconstruction (Corr = ", round(beta_corr, 3), ")"),
       subtitle = "Red line should match Black dashed line") +
  theme_bw() + theme(legend.position = "bottom")

# ------------------------------------------------------------------------------
# 4. RESULTS
# ------------------------------------------------------------------------------
grid.arrange(p1, p2, nrow = 1)

cat("\n========================================\n")
cat("      PLS REGRESSION VERIFICATION       \n")
cat("========================================\n")
cat(sprintf("1. Test R-Squared:      %.4f  [%s]\n", r2_test, ifelse(r2_test > 0.9, "PASS", "FAIL")))
cat(sprintf("2. Beta Shape Corr:     %.4f  [%s]\n", beta_corr, ifelse(beta_corr > 0.95, "PASS", "FAIL")))

if(r2_test > 0.9 && beta_corr > 0.95) {
  cat("\nCONCLUSION: ✅ PLS Works. It predicts accurately and learns the correct shape.\n")
} else {
  cat("\nCONCLUSION: ❌ Something is wrong.\n")
}
```

```{r}
# ==============================================================================
# PLS VERIFICATION: Prediction & Beta Recovery
# ==============================================================================

library(fda)
library(ggplot2)
library(gridExtra)
library(FSHybridPLS) # Your package

# ------------------------------------------------------------------------------
# 1. GENERATE TRUTH (Linear Model)
# ------------------------------------------------------------------------------
generate_regression_truth <- function(n_sample = 200, n_eval = 100, noise_sd = 0.01) {

  eval_points <- seq(0, 1, length.out = n_eval)
  basis_obj <- create.bspline.basis(c(0, 1), 20)

  # --- A. True Coefficient Function (Beta) ---
  # A distinct shape: Sine wave with a trend
  true_beta_vec <- sin(2 * pi * eval_points) + 1.5 * eval_points
  # Center it (standard practice)
  true_beta_vec <- true_beta_vec - mean(true_beta_vec)

  # --- B. Functional Predictors (X) ---
  # Smooth random curves
  coefs <- matrix(rnorm(20 * n_sample), 20, n_sample)
  # Smooth the coefficients to make curves realistic
  smooth_mat <- chol(ar1_cov(20, 0.8))
  coefs <- t(smooth_mat) %*% coefs

  fd_obj <- fd(coefs, basis_obj)
  X_vals <- eval.fd(eval_points, fd_obj) # 100 x 200

  # --- C. Response (Y) ---
  # Y = Integral( X(t) * Beta(t) ) + Noise
  # Approx integral as dot product * dt
  dt <- 1 / n_eval

  # Calculate signal
  # (100 x 200) * (100 x 1) -> broadcasting requires care or loop
  # Signal = colSums( X * beta ) * dt
  signal <- colSums(X_vals * true_beta_vec) * dt

  # Add noise
  y <- signal + rnorm(n_sample, sd = noise_sd)

  # Scalar Z (Noise only, to prove PLS ignores it)
  Z <- matrix(rnorm(n_sample), n_sample, 1)

  W <- predictor_hybrid(Z, list(fd_obj), eval_points)

  return(list(
    W = W, y = y, 
    true_beta = true_beta_vec,
    eval_points = eval_points
  ))
}

# ------------------------------------------------------------------------------
# 2. RUN SIMULATION
# ------------------------------------------------------------------------------
set.seed(42)
# Generate
sim_data <- generate_regression_truth(n_sample = 200, noise_sd = 0.01)

# Split (Train/Test)
processed <- split_and_normalize.all(sim_data$W, sim_data$y, train_ratio = 0.7)

# Fit PLS
# Using 4 components (should be enough to capture the beta shape)
# Lambda = 0 ensures we test the core algorithm, not the penalty
fit <- fit.hybridPLS(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 4, 
  lambda = c(0),
  validation_data = list(W_test = processed$predictor_test, y_test = processed$response_test)
)

# ------------------------------------------------------------------------------
# 3. VERIFICATION PLOTS
# ------------------------------------------------------------------------------

# --- A. Prediction Check (Test Set) ---
# Reconstruct Y_pred for Test Set using the Final Beta (Comp 4)
beta_final <- fit$beta[[4]]
y_pred <- inprod.predictor_hybrid(processed$predictor_test, beta_final)
y_true <- processed$response_test

r2_test <- 1 - sum((y_true - y_pred)^2) / sum((y_true - mean(y_true))^2)

df_pred <- data.frame(Observed = y_true, Predicted = y_pred)

p1 <- ggplot(df_pred, aes(x = Observed, y = Predicted)) +
  geom_point(alpha = 0.6, color = "blue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = paste0("Prediction Accuracy (R2 = ", round(r2_test, 3), ")"),
       subtitle = "Points should lie on the red diagonal") +
  theme_bw()

# --- B. Beta Recovery Check ---
# Extract estimated functional beta
beta_est_fd <- fit$beta[[4]]$functional_list[[1]]
beta_est_vec <- eval.fd(sim_data$eval_points, beta_est_fd)

# Normalize both to compare SHAPE (scale might differ due to normalization)
scale_vec <- function(x) as.vector((x - mean(x)) / sd(x))

df_beta <- data.frame(
  Time = sim_data$eval_points,
  True = scale_vec(sim_data$true_beta),
  Estimated = scale_vec(beta_est_vec)
)

beta_corr <- cor(df_beta$True, df_beta$Estimated)

p2 <- ggplot(df_beta, aes(x = Time)) +
  geom_line(aes(y = True, color = "True Beta"), linewidth = 1.2, linetype = "dashed") +
  geom_line(aes(y = Estimated, color = "PLS Estimate"), linewidth = 1) +
  scale_color_manual(values = c("True Beta" = "black", "PLS Estimate" = "red")) +
  labs(title = paste0("Beta Reconstruction (Corr = ", round(beta_corr, 3), ")"),
       subtitle = "Red line should match Black dashed line") +
  theme_bw() + theme(legend.position = "bottom")

# ------------------------------------------------------------------------------
# 4. RESULTS
# ------------------------------------------------------------------------------
grid.arrange(p1, p2, nrow = 1)

cat("\n========================================\n")
cat("      PLS REGRESSION VERIFICATION       \n")
cat("========================================\n")
cat(sprintf("1. Test R-Squared:      %.4f  [%s]\n", r2_test, ifelse(r2_test > 0.9, "PASS", "FAIL")))
cat(sprintf("2. Beta Shape Corr:     %.4f  [%s]\n", beta_corr, ifelse(beta_corr > 0.95, "PASS", "FAIL")))

if(r2_test > 0.9 && beta_corr > 0.95) {
  cat("\nCONCLUSION: ✅ PLS Works. It predicts accurately and learns the correct shape.\n")
} else {
  cat("\nCONCLUSION: ❌ Something is wrong.\n")
}
```

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

## is normalization hurtful?
```{r}
# ==============================================================================
# DIAGNOSTIC: Is Normalization Killing PLS?
# ==============================================================================

library(fda)
library(Matrix)
library(ggplot2)

# 1. GENERATE LATENT DATA (Signal defined by Variance)
# ------------------------------------------------------------------------------
# We create 2 components:
# Comp 1: Huge Variance (SD=10) -> Should be found FIRST
# Comp 2: Small Variance (SD=2) -> Should be found SECOND
generate_variance_signal <- function(n_sample = 200) {
  t1 <- rnorm(n_sample, sd = 10)
  t2 <- rnorm(n_sample, sd = 2)

  # Orthogonalize exactly
  T_mat <- svd(cbind(t1, t2))$u %*% diag(c(10, 2))

  # Functional Predictor X (Evaluation points 0 to 1)
  eval_points <- seq(0, 1, length.out=100)
  shape1 <- sin(2*pi*eval_points)
  shape2 <- cos(2*pi*eval_points)

  X <- outer(T_mat[,1], shape1) + outer(T_mat[,2], shape2) + rnorm(n_sample*100, sd=0.1)

  # Scalar Predictor Z (Just noise for now, to isolate Functional issue)
  Z <- matrix(rnorm(n_sample * 5), n_sample, 5)

  # Response Y (Driven by T1 mostly)
  y <- T_mat[,1] * 1 + T_mat[,2] * 1 + rnorm(n_sample, sd=0.1)

  # Create Objects
  basis_obj <- create.bspline.basis(c(0, 1), 20)
  fd_obj <- Data2fd(eval_points, t(X), basis_obj)

  W <- predictor_hybrid(Z, list(fd_obj), eval_points)

  return(list(W = W, y = y, true_scores = T_mat))
}

set.seed(123)
data <- generate_variance_signal(200)

# ==============================================================================
# SCENARIO A: RAW DATA (Preserve Variance)
# ==============================================================================
message("\n--- Running Scenario A: RAW DATA (Variance Preserved) ---")

# Manual Centering (Required for PLS, but NO scaling)
y_centered <- data$y - mean(data$y)
W_centered <- data$W
# (Assuming your PLS centers internally or we can pass centered data. 
#  For strictness, we assume Z is roughly centered by generation.)

fit_raw <- fit.hybridPLS(
  W = W_centered, 
  y = y_centered, 
  n_iter = 2, 
  lambda = c(0) # No penalty to isolate normalization issue
)

rho_raw <- do.call(cbind, fit_raw$rho)
cor_raw <- cor(rho_raw, data$true_scores)
print("Correlation (Estimated vs True):")
print(round(cor_raw, 3))

# ==============================================================================
# SCENARIO B: NORMALIZED DATA (Variance Destroyed)
# ==============================================================================
message("\n--- Running Scenario B: NORMALIZED DATA (Variance = 1) ---")

# Run your normalization pipeline
# We use train_ratio = 0.99 just to keep most data for the check
processed <- split_and_normalize.all(data$W, data$y, train_ratio = 0.99)

# Extract the TRUE scores corresponding to the training set indices
train_idx <- processed$details$split_indices$train
true_scores_subset <- data$true_scores[train_idx, ]

fit_norm <- fit.hybridPLS(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 2, 
  lambda = c(0)
)

rho_norm <- do.call(cbind, fit_norm$rho)
cor_norm <- cor(rho_norm, true_scores_subset)
print("Correlation (Estimated vs True):")
print(round(cor_norm, 3))

# ==============================================================================
# RESULT INTERPRETATION
# ==============================================================================
# Look at the diagonal of the correlation matrices.
# Ideally: [1,1] > 0.9 (Comp 1 found), [2,2] > 0.9 (Comp 2 found).

score_1_raw <- abs(cor_raw[1,1])
score_1_norm <- abs(cor_norm[1,1])

cat("\n========================================\n")
cat("          DIAGNOSTIC RESULTS            \n")
cat("========================================\n")
cat(paste0("Scenario A (Raw) Score 1 Recovery: ", round(score_1_raw, 3), "\n"))
cat(paste0("Scenario B (Norm) Score 1 Recovery: ", round(score_1_norm, 3), "\n"))

if (score_1_raw > 0.9 && score_1_norm < 0.9) {
  cat("\nCONCLUSION: 🚨 Normalization is the problem.\n")
  cat("By forcing unit variance, 'split_and_normalize.all' destroys the \n")
  cat("variance structure that PLS uses to identify the primary component.\n")
} else if (score_1_raw < 0.9) {
  cat("\nCONCLUSION: ⚠️  The Algorithm is the problem.\n")
  cat("Even with raw data, PLS failed to find the latent component.\n")
} else {
  cat("\nCONCLUSION: ✅ Both methods worked? (Check if simulation noise was too low)\n")
}
```

##

```{r}

# ==============================================================================
# HYBRID PLS VERIFICATION: Latent Factor & Coefficient Recovery
# ==============================================================================

library(fda)
library(ggplot2)
library(gridExtra) # For arranging plots
library(FSHybridPLS) # Your package

# ------------------------------------------------------------------------------
# 1. DATA GENERATION: The "Ground Truth"
# ------------------------------------------------------------------------------
generate_verification_data <- function(n_sample = 200, n_eval = 100) {

  eval_points <- seq(0, 1, length.out = n_eval)

  # --- A. Latent Scores (Truth) ---
  # We create 2 orthogonal components with distinct variances
  # Comp 1: High Variance (SD=10) -> The primary signal
  # Comp 2: Lower Variance (SD=4) -> The secondary signal
  set.seed(123)
  t1 <- rnorm(n_sample, sd = 10)
  t2 <- rnorm(n_sample, sd = 4)
  # Force exact orthogonality for clean testing
  T_mat <- svd(cbind(t1, t2))$u %*% diag(c(10, 4))

  # --- B. Loadings (Shapes) ---
  # Shape 1: Low frequency (Sine wave)
  # Shape 2: High frequency (Cosine wave)
  shape1 <- sin(2 * pi * eval_points)
  shape2 <- cos(4 * pi * eval_points)

  # --- C. Functional Predictor X ---
  # X = T * P' + Noise
  X_clean <- outer(T_mat[,1], shape1) + outer(T_mat[,2], shape2)
  X_noise <- matrix(rnorm(n_sample * n_eval, sd = 0.5), n_sample, n_eval)
  X <- X_clean + X_noise

  # --- D. Response Y ---
  # Y depends on both components: Y = 1*t1 - 2*t2 + Noise
  # This implies the TRUE BETA shape is roughly: (1*shape1) - (2*shape2)
  y <- T_mat[,1] * 1 + T_mat[,2] * (-2) + rnorm(n_sample, sd = 1)

  # Calculate True Beta Shape (Theoretical)
  # (Note: Normalized to SD=1 for shape comparison)
  true_beta_raw <- (1 * shape1) + (-2 * shape2)
  true_beta_norm <- (true_beta_raw - mean(true_beta_raw))/sd(true_beta_raw)

  # --- E. Objects ---
  basis_obj <- create.bspline.basis(c(0, 1), 20)
  fd_obj <- Data2fd(eval_points, t(X), basis_obj)

  # Dummy scalar (not used for signal, just to satisfy object structure)
  Z <- matrix(rnorm(n_sample), n_sample, 1)

  W <- predictor_hybrid(Z, list(fd_obj), eval_points)

  return(list(
    W = W, y = y, 
    true_scores = T_mat, 
    true_beta = true_beta_norm,
    eval_points = eval_points
  ))
}

# ------------------------------------------------------------------------------
# 2. EXECUTION
# ------------------------------------------------------------------------------
message("Generating Data...")
sim_data <- generate_verification_data(n_sample = 200)

message("Processing Data (Split & Normalize)...")
# We use a large training ratio to ensure stable estimation for verification
processed <- split_and_normalize.all(sim_data$W, sim_data$y, train_ratio = 0.8)

# Extract True Scores corresponding to the Training Set
train_idx <- processed$details$split_indices$train
true_scores_train <- sim_data$true_scores[train_idx, ]

message("Fitting Hybrid PLS...")
# Fit with 4 components (we expect the first 2 to capture the signal)
fit <- fit.hybridPLS(
  W = processed$predictor_train, 
  y = processed$response_train, 
  n_iter = 4, 
  lambda = c(0) # Small penalty for smoothness
)

>>>>

@@

  # 2. Get Scores (Rho) -- EXACT
  rho_current <- exact_inprod(W_current, xi_current)

@@

ggplot(df_long, aes(x=factor(Iteration), y=value, fill=name)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(trans="log10") +
  labs(title="Annihilation Torture Test",
       subtitle="Even with high collinearity and penalties, leakage must be ~0",
       y="Magnitude (Log10)",
       x="Iteration") +
  theme_minimal() +
  scale_fill_manual(values=c("firebrick", "steelblue")) +
  geom_hline(yintercept=1e-10, linetype="dashed")
```
```{r}
# ==============================================================================
# SIMULATION: Annihilation Audit (Using Modified fit.hybridPLS)
# ==============================================================================

@@

# ------------------------------------------------------------------------------
# 1. HELPER: EXACT INNER PRODUCT
# ------------------------------------------------------------------------------
# We use this to audit the result. It effectively replicates get_rho logic
# but ensures we control the inputs for verification.
audit_score_magnitude <- function(W, xi, J_list) {

  # 1. Scalar Part
  # (N x P) %*% (P x 1)
  scores <- as.matrix(W$Z) %*% t(as.matrix(xi$Z))

  # 2. Functional Part (Matrix Algebra)
  for(k in seq_along(W$functional_list)) {
    Theta_j <- t(W$functional_list[[k]]$coefs)
    gamma_j <- xi$functional_list[[k]]$coefs
    B_j     <- J_list[[k]]

    # Exact Integral: Theta * B * gamma
    scores <- scores + Theta_j %*% (B_j %*% gamma_j)
  }

  # Return L2 Norm (Magnitude)
  return(sqrt(sum(scores^2)))
}

# ------------------------------------------------------------------------------
# 2. DATA GENERATION
# ------------------------------------------------------------------------------
generate_torture_data <- function(n_sample = 100) {
  t1 <- rnorm(n_sample, sd=10)
  t2 <- rnorm(n_sample, sd=0.1)

  eval_points <- seq(0, 1, length.out=100)
  basis <- create.bspline.basis(c(0,1), 15)

  X1 <- outer(t1, sin(2*pi*eval_points)) 
  fd1 <- Data2fd(eval_points, t(X1), basis)

  X2 <- outer(t2, sin(10*pi*eval_points)) + matrix(rnorm(n_sample*100, sd=0.01), 100, n_sample)
  fd2 <- Data2fd(eval_points, t(X2), basis)

  Z <- matrix(rnorm(n_sample * 50), n_sample, 50)
  y <- t1 * 0.5 + t2 * 10 + rnorm(n_sample)

# ------------------------------------------------------------------------------
# 3. RUN SIMULATION
# ------------------------------------------------------------------------------
set.seed(999)
data <- generate_torture_data(100)
processed <- split_and_normalize.all(data$W, data$y, train_ratio=0.99)
W_train <- processed$predictor_train
y_train <- processed$response_train

# PRE-CALCULATE GRAM MATRICES (Required for audit helper)
basis_obj <- W_train$functional_list[[1]]$basis
J_matrix  <- fda::inprod(basis_obj, basis_obj)
gram_list_audit <- list(J_matrix, J_matrix)

# A. RUN THE FIT (One Shot)
# ------------------------------------------------------------------------------
n_comp <- 5
message("Fitting model...")

# NOTE: We attach gram_list to input W in case your internal functions use it
W_train$gram_list <- gram_list_audit

fit <- fit.hybridPLS(
  W = W_train, 
  y = y_train, 
  n_iter = n_comp, 
  lambda = c(10, 0.01)
)

# B. AUDIT THE STORED OBJECTS
# ------------------------------------------------------------------------------
message("Auditing stored W and Xi objects...")

results_df <- data.frame()

for(l in 1:n_comp) {

  # Retrieve Stored Objects directly from output
  W_before <- fit$W[[l]]      # Data used to find Component l
  W_after  <- fit$W[[l + 1]]  # Data AFTER Component l was removed (Stored in l+1)
  xi_current <- fit$xi[[l]]   # The direction found

  # Check 1: Signal Strength (Should be Large)
  # < W[l], xi[l] >
  signal_mag <- get_rho(W_before, xi_current)

  # Check 2: Annihilation / Leakage (Should be ZERO)
  # < W[l+1], xi[l] >
  leakage_mag <- get_rho(W_after, xi_current)

  # Log
  status <- ifelse(leakage_mag < 1e-2, "PASS", "FAIL")

  cat(sprintf("Comp %d: Signal=%.2e | Leakage=%.2e | %s\n", 
              l, signal_mag, leakage_mag, status))
}

# ------------------------------------------------------------------------------
# 4. VISUALIZATION
# ------------------------------------------------------------------------------
cat("\n=======================================================\n")
cat("      FIT OBJECT INTEGRITY AUDIT \n")
cat("=======================================================\n")
print(results_df)

ggplot(df_long, aes(x=factor(Iteration), y=value, fill=name)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(trans="log10") +
  labs(title="Annihilation Audit (Using fit.hybridPLS Output)",
       subtitle="Verifying orthogonality of stored W objects against stored Xi directions",
       y="Projection Magnitude (Log10)",
       x="Component") +
  theme_minimal() +
  scale_fill_manual(values=c("firebrick", "steelblue")) +
  geom_hline(yintercept=1e-10, linetype="dashed")
```
